---
title: "Homework2"
author: "Sydney Vie"
date: "2022-09-08"
output: html_document
---

# Load packages and data
## Import class one survey data from our Github site (https://github.com/kijohnson/ADA) (The dataset is called ‘Class 1    Survey Fall 2022.csv’), calling the R dataframe that you create C1survey
```{r set up, warning=FALSE}

library(tidyverse) # data wrangling
library(haven) # for SPSS, Stata, SAS, and xpt files
library(openxlsx) # for excel

C1survey <- read_csv("https://raw.githubusercontent.com/kijohnson/ADA/master/Class%201%20Survey%20Fall%202022.csv")
# C1survey is a data set including information from a survey given to students of ADA in Fall 2022 before their first class.

```

## Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of     variables in the dataframe.
```{r describe dataset}

dim(C1survey)
# The C1survey dataset has 42 observations and 27 variables.



```

# Data wrangling
## a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.
```{r rename}

names(C1survey)[1:27] <- c("ID", "like_cats", "like_dogs", "desert_island_item", 
                           "life_slogan", "favorite_day", "bird_type",
                           "favorite_food", "favorite_drink", "favorite_season",
                           "favorite_month", "hobby", "degree_program",
                           "primary_specialization", "primary_software",
                           "R_experience_level", "coding_comfort_level", 
                           "coding_experience_time", "three_learning_hopes", 
                           "public_health_interest", "favorite_number", 
                           "birth_day", "birth_month", "home_country",
                           "home_state", "home_city", "highest_education_level") 
# rename columns to be shorter and more descriptive of the variable

colnames(C1survey)
# display new column names


```

## Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.
```{r variable types}

table(sapply(C1survey, class))
# 23 variables are character variables
# 4 variables are double variables


```

## Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r birth day and birth month variables}

C1survey$birth_day
# check values for birth day
# there are 2 strange values - "June 8th" and "13-Aug-97"

C1survey$birth_month
# check values for birth months
# there are four strange values

C1survey$birth_day <- ifelse(C1survey$birth_day == "June 8th", "8", C1survey$birth_day)
C1survey$birth_day <- ifelse(C1survey$birth_day == "13-Aug-97", "13", C1survey$birth_day)
# recode these variables into integers

C1survey$birth_day
# check values again

C1survey$birth_day <- as.numeric(C1survey$birth_day)
# convert character class to numeric


C1survey$birth_month[C1survey$birth_month=="June 8th, 1998"]<-"6"
C1survey$birth_month[C1survey$birth_month=="October"]<-"10"
C1survey$birth_month[C1survey$birth_month=="August"]<-"8"
C1survey$birth_month[C1survey$birth_month=="March"]<-"3"
# recode these character values into integers

C1survey$birth_month
# check values again

C1survey$birth_month <- as.numeric(C1survey$birth_month)
# convert birth day and birth month from character to numeric class

class(C1survey$birth_month)
class(C1survey$birth_day)
# check classes

summary(C1survey$birth_day)
# get summary statistics for the birth day variable
# minimum is 1 and maximum is 30
# there is 1 NA

summary(C1survey$birth_month)
# get summary statistics for birth month variable
# minimum is 1 and maximum is 12
# there is 1 NA

C1survey <- C1survey %>% 
  drop_na(c(birth_day, birth_month))
# remove rows which have missing values for birth day or birth month
# there are now 41 observations

median(C1survey$birth_day)
# the median birth day is 15

median(C1survey$birth_month)
# the median birth month is July

```


## Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter=12/1 to 2/29, spring=3/1 to 5/31, summer=6/1 to 8/31, fall=9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.
```{r birth season}

C1survey <- C1survey %>% 
  mutate(birth_season = case_when(birth_month %in% c(12, 1, 2) ~ "Winter",
                                  birth_month %in% c(3,4,5) ~ "Spring",
                                  birth_month %in% c(6,7,8) ~ "Summer",
                                  birth_month %in% c(9,10,11) ~ "Fall"))
# create a variable for the seasons in which respondents were born


seasonTable <- with(C1survey, table(birth_month, birth_season))
print(seasonTable)
# check coding is correct
# looks good

?addmargins
# get help on the arguments in the addmargins function

addmargins(seasonTable, FUN = sum)
# determine how many classmates were born in each season
# 9 born in winter, 8 born in spring, 13 born in summer, 11 born in fall


```

## Pick your favorite variable to analyze, come up with a question you want to answer with that variable, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).
```{r create and answer question}

# Is there an association between birth season and favorite season?

C1survey$favorite_season <- as.factor(C1survey$favorite_season)
# convert favorite season from a character to a factor variable

C1survey$birth_season <- as.factor(C1survey$birth_season)

class(C1survey$favorite_season)
class(C1survey$birth_season)
# check that it is a factor

chisq.test(table(C1survey$birth_season, C1survey$favorite_season))
# perform a chi-square test as an omnibus test to see if there is an association between the two variables
# expected cell counts less than 5, use Fisher's exact test

fisher.test(table(C1survey$birth_season, C1survey$favorite_season))
# perform a fisher's exact test
# p = 0.311, there is not a significant relationship between birth season and favorite season

```


